Space astrometry of the very massive $\sim$$150\,M_\odot$ candidate runaway star VFTS682

### 2018-08-22 MNRASformat/VFTS682_MNRASL.tex
ready to submit to MNRASL

How very massive stars form is an intriguing open question in
stellar astrophysics.  VFTS682 is among the most massive stars known,
with an inferred initial mass of $\gtrsim$$150\,M_\odot$. It is located
in 30 Doradus at a projected distance of 29\,pc from the central
cluster R136. % Its apparent isolation
The absence of other massive stars in its immediate vicinity
led to two intriguing hypotheses: % posed in earlier work:
either it formed in relative isolation % through a new mode of star formation
or it was ejected dynamically from the central cluster. 
In this context, 
we investigate the kinematics of
VFTS682 as obtained by \emph{Gaia} and multi-epoch \emph{Hubble Space
  Telescope} astrometry. We derive a projected velocity relative to
the cluster of $38\pm17\kms$ ($1\sigma$ confidence interval). Although
the error bars are substantial, two completely independent measures
suggest that VFTS682 might be a slow runaway
ejected from the central cluster, given its distance and the inferred
age. This hypothesis is further supported by a variety of
circumstantial clues. The central cluster is known to harbor several
other stars more massive than $150\,M_\odot$ of similar spectral
type. The very massive O stars VFTS16 (and possibly VFTS72), which has
recently been proposed as a dynamically ejected runaway, provides
direct evidence that R136 is indeed capable of ejecting some of its
most massive members, consistent with the predictions of numerical
simulations. If future data confirm the runaway nature, this would make VFTS682
the most massive runaway star known to date. 
\end{abstract}

@@

 % \begin{acknowledgements}
{   \small
   We are grateful to J.~Heyl, M.\,C.\,Ramirez-Tannus, S.~N.~Shore, and S.\,Torres
   % and C.\,J.\,Evans
   for help and discussions. We also thank the HSTPROMO
(High-resolution Space Telescope PROper MOtion) Collaboration for
sharing and discussing ideas relevant to this work. SdM acknowledges
the European Unions Horizon 2020 research and innovation programme
from the European Research Council (ERC), Grant agreement No. 715063
and VHB the NRC-Canada Plaskett Fellowship. This work has made use of data from the European Space Agency (ESA) mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia} Data Processing and Analysis Consortium (DPAC, \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the DPAC has been provided by national institutions, in particular the institutions
participating in the {\it Gaia} Multilateral Agreement. 
% \end{acknowledgements}
}

>>>>

How very massive stars form is an intriguing open question in
stellar astrophysics.  VFTS682 is among the most massive stars known,
with an inferred initial mass of $\gtrsim$$150\,M_\odot$. It is located
in 30 Doradus at a projected distance of 29\,pc from the central
cluster R136. 
Its apparent isolation led to two intriguing hypotheses: 
either it formed in relative isolation 
or it was ejected dynamically from the cluster. 
We investigate the kinematics of
VFTS682 as obtained by \emph{Gaia} and multi-epoch \emph{Hubble Space Telescope} astrometry. We derive a projected velocity relative to
the cluster of $38\pm17\kms$ ($1\sigma$ confidence interval). Although
the error bars are substantial, two independent measures
suggest that VFTS682 might be a runaway
ejected from the central cluster. This hypothesis is further supported by a variety of
circumstantial clues. The central cluster is known to harbor 
other stars more massive than $150\,M_\odot$ of similar spectral
type. The very massive O stars VFTS16 (and possibly VFTS72), 
recently 
proposed as a dynamically ejected runaway, provides
direct evidence that R136 is
capable of ejecting some of its
most massive members, consistent with the predictions of numerical
simulations. If future data confirm the runaway nature, this would make VFTS682
the most massive runaway star known to date. 
\end{abstract}

@@

 % \begin{acknowledgements}
{   \small
   We are grateful to P.~Crowther, J.~Heyl, M.\,C.\,Ramirez-Tannus, S.~N.~Shore, and S.\,Torres
   % and C.\,J.\,Evans
   for help and discussions. We also thank the HSTPROMO
(High-resolution Space Telescope PROper MOtion) Collaboration for
sharing and discussing ideas relevant to this work. SdM acknowledges
the European Unions Horizon 2020 research and innovation programme
from the European Research Council (ERC), Grant agreement No. 715063
and VHB the NRC-Canada Plaskett Fellowship. This work has made use of data from the European Space Agency (ESA) mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia} Data Processing and Analysis Consortium (DPAC, \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the DPAC has been provided by national institutions, in particular the institutions
participating in the {\it Gaia} Multilateral Agreement. 
% \end{acknowledgements}
}
